\Title

### 2020-12-15 manuscript/manuscript.tex

\documentclass[twocolumn]{article}
\usepackage[left=0.7in,right=0.7in,top=1in,bottom=1in]{geometry}
\setlength{\columnsep}{2\columnsep}
\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[longend,linesnumbered]{algorithm2e}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[round,authoryear,sort]{natbib}
\usepackage{url}
\usepackage[pdftex,colorlinks=true]{hyperref}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{setspace}

@@

\section{The equivalent sources technique}

The Green's equivalent layer theorem states that any harmonic potential field
generated by any three dimensional physical-property distribution can by
reproduced by a continuous two dimensional surface distribution.
This introduces an ambiguity on the distribution that generates
a potential field: multiple physical-property distributions can generate the
same potential field.
This represents an obstacle for determining the true shape and
physical-property distribution that generates any measured potential field.
Nevertheless this ambiguity can also be exploited for developing pre-processing
techniques like interpolating potential field data.

Approximating the continuous distribution by a finite and discrete density
distribution is the basis of the \emph{equivalent sources technique}
\citep{dampney1969}.
The most common approach is to make use of a set of point sources (the
\emph{equivalent sources}) located beneath the observation points.

A classical approach for applying the equivalent sources technique would be to
propose a set of point sources that generate the same harmonic field that has
been measured.
Lets consider a set of gravity acceleration measurements $g_z$ on points
$\{\mathbf{p}_1, \ldots, \mathbf{p}_N\}$ and a set of point sources with mass
$\{m_1, \ldots, m_M\}$ located on points
$\{\mathbf{q}_1, \ldots, \mathbf{q}_M\}$.
The point sources must generate the same gravity field as the one measured:

\begin{equation}
    g_z(\mathbf{p}) = \mathbf{A} \mathbf{m},
    \label{eq:linear-problem-g}
\end{equation}

\noindent where $g_z(\mathbf{p})$ is the vector containing the measured
field on the points $\mathbf{p}$, $\mathbf{m}$ is the vector containing the
masses of the point sources and $\mathbf{A}$ a matrix whose
elements can be obtained as the vertical component of the gravitational
acceleration measured on $\mathbf{p}$ due to the point mass on $\mathbf{q}$:

\begin{equation}
    a_{ij} = G \frac{(z_i - z_j)}{|\mathbf{p}_i - \mathbf{q}_j|^3},
\end{equation}

\noindent where
$G = \SI{6.674e-11}{\cubic\meter\per\kilogram\per\second\squared}$ is the
universal gravitational constant, $z_i$ and $z_j$ are the vertical components
of the $\mathbf{p}_i$ and $\mathbf{q}_j$ points, respectively.

By solving equation~\ref{eq:linear-problem-g} we can obtain the values of
$\mathbf{m}$ that better reproduce the measured field.
This source distribution along with the fitted masses can be used to predict
gravity values on any point where the potential field has not been measured.

\subsection{Generalized equivalent sources}

It's possible to generalize the equivalent sources technique to any harmonic
field through a Green's function problem, regardless the nature of the field.

By definition, any harmonic field $V$ must satisfy the Laplace's equation:

\begin{equation}
    \nabla^2 V = 0.
\end{equation}

Considering as boundary condition that $V$ is equal to zero at infinity, then
its fundamental solution $\psi$ (also known as Green's function or radial basis
function) must satisfy \citep[p.~210]{vladimirov1979}:

\begin{equation}
    \nabla_\mathbf{p}^2 \psi(\mathbf{p}, \mathbf{q}) =
        - 4\pi \delta(\mathbf{p}, \mathbf{q}),
\end{equation}

\noindent where $\delta(\mathbf{p}, \mathbf{q})$ is the Dirac's delta function
and $\nabla_\mathbf{p}$ symbolize the Nabla operator applied on the coordinates
$\mathbf{p}$.

The fundamental solution of any harmonic field can be obtained as follows
\citep[p.~37]{blakely1995}:

\begin{equation}
    \psi(\mathbf{p}, \mathbf{q})
    =
    \int
    \frac{\delta(\mathbf{p}, \mathbf{q})}{|\mathbf{p} - \mathbf{q}|}
    \textrm{d}v
    =
    \frac{1}{|\mathbf{p} - \mathbf{q}|}.
\end{equation}

This shows that the Green's function associated with any harmonic field is
equal to the inverse of the Euclidean distance between the evaluation and the
source points.
Thus any harmonic field $V$ can be expressed as \citep[p.~37]{blakely1995}:

\begin{equation}
    V(\mathbf{p})
    =
    \int \psi(\mathbf{p}, \mathbf{q}) \rho(\mathbf{q}) d\mathbf{q}
    =
    \int \frac{\rho(\mathbf{q})}{|\mathbf{p} - \mathbf{q}|}  d\mathbf{q},
\end{equation}

\noindent where $\rho(\mathbf{q})$ is the density distribution of a certain
physical-property evaluated on the point $\mathbf{q}$. Following
\citet{dampney1969}, this density distribution can be approximated by a finite
and discrete set of point sources located on $\{\mathbf{q}_1, \ldots,
\mathbf{q}_M\}$:

\begin{equation}
    \rho(\mathbf{q}) =
        \sum\limits_{j=1}^{M} c_j \delta(\mathbf{q} - \mathbf{q_j}),
\end{equation}

\noindent where $c_j$ is a coefficient related to the point source located on
$\mathbf{q}_j$, with the same units as the physical-property.
Then the harmonic field $V$ can be approximated by:

\begin{equation}
    V(\mathbf{p})
    =
    \sum\limits_{j=1}^{M} \frac{c_j}{|\mathbf{p} - \mathbf{q}_j|}.
    \label{eq:eql-field}
\end{equation}

In case we have measured the value of the harmonic field on points
$\{\mathbf{p}_1, \ldots, \mathbf{p}_N\}$, we can adjust the values of the $c$
coefficients so they satisfy:

\begin{equation}
    V(\mathbf{p}_i)
    =
    \sum\limits_{j=1}^{M} \frac{c_j}{|\mathbf{p}_i - \mathbf{q}_j|}
    \quad \forall i=1,\ldots,N.
\end{equation}

This system of linear equations can be written in an more algebraic way:

\begin{equation}
    \mathbf{d} = \mathbf{A} \mathbf{c}
    \label{eq:linear-problem}
\end{equation}

\noindent where $\mathbf{d}$ is a vertical vector containing the measured
values of $V$ on the observation points, $\mathbf{c}$ a vertical vector
containing the $c$ coefficients and $\mathbf{A}$ is the \emph{Jacobian matrix},
whose elements are defined as the Green's function evaluated on each pair of
points $\mathbf{p}$ and $\mathbf{q}$:

\begin{equation}
    a_{ij}
    =
    \psi(\mathbf{p}_i, \mathbf{q}_j)
    =
    \frac{1}{|\mathbf{p}_i - \mathbf{q}_j|}
\end{equation}

By solving equation~\ref{eq:linear-problem} we can obtain the values of
$\mathbf{c}$ that better reproduce the measured field.
Then they can be used to predict the value of the harmonic field on any other
point where no measurement has been taken by evaluating
equation~\ref{eq:eql-field}.

\subsection{Fitting source coefficients}
\label{sec:eql-inversion}

Solving equations~\ref{eq:linear-problem-g} and~\ref{eq:linear-problem} is
equivalent to performing an inversion and thus solving an ill-posed problem by
nature due to the non-uniqueness and the instability of the possible solutions.
Therefore is usually needed to add regularization functions in order to obtain
stable solutions
(\citealp[p.~52]{menke1989}; \citealp{cooper2000}; \citealp{oliveira2013}).

We can obtain the values of the source coefficients $\mathbf{c}$ that better
reproduce the measured field by minimizing an objective or misfit function
$\phi(\mathbf{c})$ through a least squares method:

\begin{equation}
    \phi(\mathbf{c}) =
        \left\lVert
            \mathbf{d} - \mathbf{A}\mathbf{c}
        \right\rVert ^ 2
        + \lambda_d \left\lVert \mathbf{c} \right\rVert ^2,
    \label{eq:misfit-unscaled}
\end{equation}

\noindent where $\left\lVert \cdot \right\rVert ^ 2$ symbolizes the L$_2$ norm
and $\lambda_d$ is a \emph{damping} parameter with the same units as the
Jacobian matrix elements.
The second term of the equation applies a zeroth-order Tikhonov
\citep{tikhonov1977} regularization to stabilize the solution.

In order to simplify the choice of the \emph{damping} parameter we scale the
Jacobian matrix so each column has unit variance.
We define a diagonal matrix $\mathbf{S}$, whose elements corresponds to the
standard deviation of each column of the Jacobian matrix $\mathbf{A}$:

\begin{equation}
    \mathbf{S} =
    \begin{bmatrix}
        \sigma_1 & \cdots &0 \\
        \vdots & \ddots & \vdots \\
        0 & \cdots & \sigma_M
    \end{bmatrix}
    ,
\end{equation}

\noindent so equation~\ref{eq:linear-problem} can be written as:

\begin{equation}
    \mathbf{d}
    =
    \mathbf{A} \mathbf{S}\inv \mathbf{S} \mathbf{c}
    =
    \left[
        \mathbf{A} \mathbf{S}\inv
    \right]
    \left[
        \mathbf{S} \mathbf{c}
    \right]
    =
    \hat{\mathbf{A}} \hat{\mathbf{c}}
\end{equation}

\noindent where $\hat{\mathbf{A}}$ is the scaled Jacobian matrix and
$\hat{\mathbf{c}}$ is a vector containing scaled coefficients.

The misfit function defined on equation~\ref{eq:misfit-unscaled} can be then
rewritten as follows:

\begin{equation}
    \phi(\hat{\mathbf{c}}) =
        \left\lVert \mathbf{d}
        - \hat{\mathbf{A}}\hat{\mathbf{c}} \right\rVert ^ 2
        + \lambda \left\lVert \hat{\mathbf{c}} \right\rVert ^2,
    \label{eq:misfit}
\end{equation}

\noindent where now $\lambda$ is a \emph{dimensionless} damping parameter.

Misfit function from equation~\ref{eq:misfit} is minimized through Cholesky
decomposition.
If the number of observation points is equal or larger to the number of
sources, then the linear problem defined on equation~\ref{eq:linear-problem} is
well- or over-determined, respectively. Therefore, the scaled coefficients that
minimize the misfit function can be obtained through \citep{menke1989}:

\begin{equation}
    \hat{\mathbf{c}} =
        \left[
            \hat{\mathbf{A}}^\text{T} \hat{\mathbf{A}}
            + \lambda^2 \mathbf{I}
        \right]\inv
        \hat{\mathbf{A}}^\text{T}
        \mathbf{d}.
\end{equation}

On the other hand, if there are more sources than observation points, the
problem is under-determined and the scaled coefficients can be obtained as
\citep{menke1989}:

\begin{equation}
    \hat{\mathbf{c}} =
    \hat{\mathbf{A}}^\text{T}
        \left[
            \hat{\mathbf{A}} \hat{\mathbf{A}}^\text{T}
            + \lambda^2 \mathbf{I}
        \right]\inv
        \mathbf{d}.
\end{equation}

Once the scaled coefficients are obtained, the sources' coefficients
$\mathbf{c}$ can be obtained by rescaling
$\hat{\mathbf{c}}$:

\begin{equation}
    \mathbf{c} = \mathbf{S}\inv \hat{\mathbf{c}}.
\end{equation}

>>>>

\documentclass[twocolumn]{article}
%\usepackage[left=0.7in,right=0.7in,top=1in,bottom=1in]{geometry}
\setlength{\columnsep}{2\columnsep}
\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[longend,linesnumbered]{algorithm2e}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[round,authoryear,sort]{natbib}
\usepackage{url}
\usepackage[pdftex,colorlinks=true]{hyperref}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{setspace}

@@

\section{Methodology}

\subsection{The equivalent sources technique}

Here, we will follow the ``generalized equivalent sources'' of
\citet{cordell1992} and assume that any harmonic function $d(\mathbf{p})$ can
be approximated by a sum of $M$ discrete point source effects

\begin{equation}
    d(\mathbf{p})
    =
    \sum\limits_{j=1}^{M} \frac{c_j}{\left\lVert \mathbf{p} - \mathbf{q}_j
    \right\rVert} \ ,
    \label{eq:eql-forward}
\end{equation}

\noindent in which
$\mathbf{p}$ and $\mathbf{q}_j$ are position vectors in a 3D Cartesian space,
$c_j$ is a scalar coefficient related to the point source located at
$\mathbf{q}_j$,
and $\lVert \cdot \rVert$ represents the $\text{L}_2$ norm.

In case we have measured the value of the harmonic field on $N$ points
$\{\mathbf{p}_1\ \mathbf{p}_2\ \ldots\ \mathbf{p}_N\}$,
we can write a set of $N$ equations of the form

\begin{equation}
    d_i
    =
    \sum\limits_{j=1}^{M} \frac{c_j}{\left\lVert \mathbf{p}_i - \mathbf{q}_j
    \right\rVert}
    \quad \forall i=1,2,\ldots,N
    \ ,
\end{equation}

\noindent where $d_i$ is the value of the potential field at the observation point $\mathbf{p}_i$.
These equations can also be expressed as

\begin{equation}
    \mathbf{d} = \mathbf{A} \mathbf{c} \ ,
    \label{eq:linear-problem}
\end{equation}

\noindent where $\mathbf{d}$ is a column vector containing the $N$ predicted
values of the observed field at the observation points,
$\mathbf{c}$ is a column vector containing the $M$ coefficients $c_j$,
and $\mathbf{A}$ is the $N \times M$ Jacobian matrix,
whose elements are

\begin{equation}
    a_{ij} = \frac{1}{\left\lVert\mathbf{p}_i - \mathbf{q}_j\right\rVert}
\end{equation}

For a given set of $N$ observed field values $\mathbf{d}^o$,
we can find a least-squares solution to
equation~\ref{eq:linear-problem} and obtain the values of
$\mathbf{c}$ that best fit the observations.
These values can, in turn, be used to predict the value of the harmonic field
at any other point outside of the sources by evaluating
equation~\ref{eq:eql-forward}.
Gridding and upward continuation can thus be achieved by predicting values on
points that fall on a regular grid or at different heights.

\subsection{Damped least-squares solution}
\label{sec:eql-inversion}

We can obtain the values of the source coefficients $\mathbf{c}$ that best
fit the observed field values $\mathbf{d}^o$ by minimizing the goal function

\begin{equation}
    \phi(\mathbf{c}) =
    \left[\mathbf{d}^o - \mathbf{A}\mathbf{c}\right]\trans
    \mathbf{W}
    \left[\mathbf{d}^o - \mathbf{A}\mathbf{c}\right]
    + \lambda_d\ \mathbf{c}\trans\mathbf{c}
    \ ,
    \label{eq:misfit-unscaled}
\end{equation}

\noindent where
$\mathbf{W}$ is a $N \times N$ diagonal matrix of data weights and
$\lambda_d$ is a positive \emph{damping} parameter with the same units as the
Jacobian matrix elements.
The second term on the right-hand side of the equation is a zeroth-order
Tikhonov regularization \citep{tikhonov1977} (also known as a damping
regularization or ridge regression) that is used to stabilize the solution.

The range of acceptable values for the damping parameter $\lambda_d$ will
depend on the values of the Jacobian matrix $\mathbf{A}$ and the coefficients.
Consequently, this range will vary (often dramatically) between datasets,
making it difficult to choose an appropriate value for $\lambda_d$ in practice.

To solve this issue, we first scale the Jacobian matrix so that its elements
are unitless and each column has unit variance.
We define a diagonal matrix $\mathbf{S}$

\begin{equation}
    \mathbf{S} =
    \begin{bmatrix}
      \sigma_1 & 0 & \cdots &0 \\
      0 & \sigma_2 & \cdots &0 \\
      \vdots & \vdots & \ddots & \vdots \\
      0  & 0 & \cdots & \sigma_M
    \end{bmatrix}_{M \times M}
    ,
\end{equation}

\noindent in which $\sigma_j$ is the standard deviation of the $j$-th column of
$\mathbf{A}$.
We then write the forward problem in equation~\ref{eq:linear-problem} as

\begin{equation}
    \mathbf{d}
    =
    \mathbf{A} \mathbf{S}\inv \mathbf{S} \mathbf{c}
    =
    \left[
        \mathbf{A} \mathbf{S}\inv
    \right]
    \left[
        \mathbf{S} \mathbf{c}
    \right]
    =
    \mathbf{B} \mathbf{m}
\end{equation}

\noindent where $\mathbf{B}$ is the scaled and unitless Jacobian matrix and
$\mathbf{m}$ is a vector containing scaled coefficients with the same units as
the data.

The goal function defined in equation~\ref{eq:misfit-unscaled} can be
rewritten as

\begin{equation}
    \phi(\mathbf{m}) =
    \left[\mathbf{d}^o - \mathbf{B}\mathbf{m}\right]\trans
    \mathbf{W}
    \left[\mathbf{d}^o - \mathbf{B}\mathbf{m}\right]
    + \lambda\ \mathbf{m}\trans\mathbf{m}
    \ ,
    \label{eq:misfit}
\end{equation}

\noindent where $\lambda$ is a \emph{dimensionless} damping parameter and
regularization is applied on the scaled coefficients $\mathbf{m}$ instead of
$\mathbf{c}$.

The vector of scaled coefficients $\hat{\mathbf{m}}$ that minimizes the goal
function can be found by solving the \emph{normal equation system}
\citep{menke1989}

\begin{equation}
    \left[
      \mathbf{B}\trans \mathbf{W} \mathbf{B} + \lambda \mathbf{I}
    \right]
    \hat{\mathbf{m}} =
    \mathbf{B}\trans\mathbf{W}
    \mathbf{d}^o.
\end{equation}

Once the scaled coefficients are obtained, the estimated unscaled coefficients
$\hat{\mathbf{c}}$ can be calculated by removing the scaling factor

\begin{equation}
    \hat{\mathbf{c}} = \mathbf{S}\inv \hat{\mathbf{m}} \ .
\end{equation}

\noindent The forward modeling operations used to perform predictions
(e.g., for interpolation and upward continuation) are left unchanged by
using vector $\hat{\mathbf{c}}$ instead of $\hat{\mathbf{m}}$.
